### 2014-11-03 thesis/litreview/litreview.tex
Small updates to litreview

\section{Statistical Methods}
\subsection{Regression}
As we stated earlier, logistic regression is a commonly-used modelling
technique for prediction in the medical domain \citep{Tu1996}, and especially
in the trauma data analysis \citep{Gabbe2005}. A series of
predictor (independent) variables (or features) determine the probability of a
particular outcome by an equation of the form
\begin{equation}
\mathrm{log}\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1X_1 + \beta_2X_2 + \dots + \beta_nX_n
\end{equation}
where $p$ is the probability of the desired outcome, $\beta_i$ are coefficients
associated with each variable, and $X_i$ are the values of the independent
variables \citep{Hosmer1989}. A stepwise selection algorithm is often used to
choose which variables to include in the model out of all available variables:
different algorithms produce different models which are then compared against
one another to find the best model. Usually, a standard maximum-likelihood
procedure is followed in order to obtain values for the $\beta_i$'s, and models
can be tested using measures such as the area under the receiver operating
characteristic (ROC) curve (AUC) and the Hosmer-Lemeshow statistic.

@@

Yang et al. \citep{Yang2010} used a linear regression model as the baseline for
performance benchmarking when trying to predict the length of stay of burn
patients. Interestingly, unlike the logistic regression models discussed above,
they did not perform any feature selection when constructing their models and
found that this gave better performance against models constructed using
correlation-based feature selection\footnote{Abbreviated CFS, correlation-based
feature selection is a technique of choosing which independent variables to
include in a model based on the hypothesis that good variable subsets should
be highly correlated with the dependent variable but not with each
other \citep{Hall2000}.}. Unlike other studies into length of stay prediction,
they also looked at predicting the length of stay at various stages of the
clinical process, rather than simply at admission. By making predictions
at the admission, acute, and post-treatment stages, they were able to decrease
their predictions' mean absolute error (MAE) from 9.532 days at admission, to
6.314 days at post-treatment, where MAE is defined as
\begin{equation}
MAE = \frac{\sum_{i=1}^n |\mathrm{PredictedLOS}_i-\mathrm{ActualLOS}_i|}{n}
\end{equation}
and PredictedLOS$_i$ is the predicted length of stay for the $i$-th instance,
ActualLOS$_i$ is the actual length of stay for the $i$-th instance, and $n$
is the total number of instances. Their approach was very much appropriate to
their medical domain (burn patients) as they tended to have longer lengths of
stay (which is more suitable for linear regression rather than logistic
regression) and the extended period of stay means that there will be enough
data from several phases of the clinical process for such patients. It is
difficult to apply such a phase method to our problem of trauma length of
stay prediction, however, as there is no standard clinical process due to
the varying nature of trauma causes and the lengths of stay are typically not
as long.

\subsection{Markov Models}
There have been a number of studies attempting to model the length of stay
using discrete-time Markov processes. Kapadia et al. constructed a transition
probability matrix using 4 states for a pediatric intensive care unit (ICU)
application \citep{Kapadia2000}:
\begin{center}
\begin{tabular}{l|llll}
& Low & Medium & High & Exit \\
\hline
Low & 0.309 & 0.223 & 0.024 & 0.444 \\
Medium & 0.137 & 0.513 & 0.096 & 0.254 \\
High & 0.086 & 0.364 & 0.379 & 0.171 \\
Exit & 0 & 0 & 0 & 1 \\
\end{tabular}
\end{center}
From the matrix, we can say that the probability of transitioning from state
`Medium' to state `Low' is 0.137. Note that each row sums to 1, and that `Exit'
is an absorbing state: once entered, one can only re-enter the `Exit' state
with 100\% certainty. This represents patients leaving the ICU. How did they
obtain these probabilities? From their data set, they classified each patient
with an appropriate state (Low, Medium, High) corresponding to the severity of
their illness based on a particular mortality score (PRISM), and did this daily
over a few months; this gave them patient records that were simply strings of
L, M and H. To calculate the probability of transitioning from `Medium' to
`Low', they counted all occurrences of `Medium' to `Low' transitions and
divided this by the number of all transitions out of `Medium', and repeated
this for every transition. Using the theory of Markov processes, they are able
to calculate the average time taken before a person enters the `Exit' state,
which is precisely the length of stay. They found that the Markov model fitted
the data well and produced predictions that were very close to the observed
results in the test set \citep{Kapadia2000}, indicating that using such models
justifies further investigation.

@@

\subsection{Artificial Neural Networks}
Artificial neural networks have been widely used in clinical prediction
following their use in diagnosing acute myocardial infarction and various
applications in radiology \citep{Baxt1995}.
One of the earliest applications of neural networks to
length of stay prediction was by Tu \& Guerriere, who used a neural network
with the following structure \citep{Tu1993} to predict length of stay in
intensive care following cardiac surgery:
\begin{figure}[h]
\caption{A graphical representation of a neural network with one hidden layer,
used by Tu \& Guerriere in predicting LOS.}
\centering
\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

@@

Yang et al. investigated the performance of linear regression, M5 (another
tree-based classification method), and support vector machines in predicting
length of stay for burn patients. The difference between their tree method,
M5, with that of Harper (CART), is that the nodes contain linear regression
functions instead of the usual discrete (two-class) outputs that are used to
successively split the data set. Additionally, they also use support vector
machines (SVMs), perhaps the most powerful classification algorithm in terms of
predictive accuracy \citep{Bellazzi2008}. SVMs work on the principle that when
the data is represented in a high enough dimensional space, a linear hyperplane
can be found where the margin between two different records is as large as
possible. In Figure \ref{fig:svm-example} from Yoo et al. \citep{Yoo2012},
hyperplane 2 is the better plane because its margin is as large as possible. 
\begin{figure}[h]
\caption{An example of finding the largest margin hyperplane for a two-class
support vector machine classifier.}
\label{fig:svm-example}
\centering
\includegraphics[scale=1.05]{images/a2-svm}
\end{figure}

>>>>

\section{Statistical Methods}
\subsection{Regression}
As we stated earlier, logistic regression is a commonly-used modelling
technique for prediction in the medical domain \citep{Tu1996}, and especially
in the trauma data analysis \citep{Gabbe2005}. A series of
predictor (independent) variables (or features) determine the probability of a
particular outcome by an equation of the form
\begin{equation*}
\mathrm{log}\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1X_1 + \beta_2X_2 + \dots + \beta_nX_n
\end{equation*}
where $p$ is the probability of the desired outcome, $\beta_i$ are coefficients
associated with each variable, and $X_i$ are the values of the independent
variables \citep{Hosmer1989}. A stepwise selection algorithm is often used to
choose which variables to include in the model out of all available variables:
different algorithms produce different models which are then compared against
one another to find the best model. Usually, a standard maximum-likelihood
procedure is followed in order to obtain values for the $\beta_i$'s, and models
can be tested using measures such as the area under the receiver operating
characteristic (ROC) curve (AUC or AUROC) and the Hosmer-Lemeshow statistic.

@@

Yang et al. \citep{Yang2010} used a linear regression model as the baseline for
performance benchmarking when trying to predict the length of stay of burn
patients. Interestingly, unlike the logistic regression models discussed above,
they did not perform any feature selection when constructing their models and
found that this gave better performance against models constructed using
correlation-based feature selection\footnote{Abbreviated CFS, correlation-based
feature selection is a technique of choosing which independent variables to
include in a model based on the hypothesis that good variable subsets should
be highly correlated with the dependent variable but not with each
other \citep{Hall2000}.}. Unlike other studies into length of stay prediction,
they also looked at predicting the length of stay at various stages of the
clinical process, rather than simply at admission. By making predictions
at the admission, acute, and post-treatment stages, they were able to decrease
their predictions' mean absolute error (MAE) from 9.532 days at admission, to
6.314 days at post-treatment, where MAE is defined as
\begin{equation*}
MAE = \dfrac{1}{n}\sum_{i=1}^n |\mathrm{PredictedLOS}_i-\mathrm{ActualLOS}_i|
\end{equation*}
and PredictedLOS$_i$ is the predicted length of stay for the $i$-th instance,
ActualLOS$_i$ is the actual length of stay for the $i$-th instance, and $n$
is the total number of instances. Their approach was very much appropriate to
their medical domain (burn patients) as they tended to have longer lengths of
stay (which is more suitable for linear regression rather than logistic
regression) and the extended period of stay means that there will be enough
data from several phases of the clinical process for such patients. It is
difficult to apply such a phase method to our problem of trauma length of
stay prediction, however, as there is no standard clinical process due to
the varying nature of trauma causes and the lengths of stay are typically not
as long.

\subsection{Markov Models}
There have been a number of studies attempting to model the length of stay
using discrete-time Markov processes. Kapadia et al. constructed a transition
probability matrix using 4 states for a pediatric intensive care unit (ICU)
application \citep{Kapadia2000}:

\begin{table}[h]
\begin{tabular}{l|llll}
& Low & Medium & High & Exit \\
\hline
Low & 0.309 & 0.223 & 0.024 & 0.444 \\
Medium & 0.137 & 0.513 & 0.096 & 0.254 \\
High & 0.086 & 0.364 & 0.379 & 0.171 \\
Exit & 0 & 0 & 0 & 1 \\
\end{tabular}
\caption{}
\end{table}
From the matrix, we can say that the probability of transitioning from state
`Medium' to state `Low' is 0.137. Note that each row sums to 1, and that `Exit'
is an absorbing state: once entered, one can only re-enter the `Exit' state
with 100\% certainty. This represents patients leaving the ICU. How did they
obtain these probabilities? From their data set, they classified each patient
with an appropriate state (Low, Medium, High) corresponding to the severity of
their illness based on a particular mortality score (PRISM), and did this daily
over a few months; this gave them patient records that were simply strings of
L, M and H. To calculate the probability of transitioning from `Medium' to
`Low', they counted all occurrences of `Medium' to `Low' transitions and
divided this by the number of all transitions out of `Medium', and repeated
this for every transition. Using the theory of Markov processes, they are able
to calculate the average time taken before a person enters the `Exit' state,
which is precisely the length of stay. They found that the Markov model fitted
the data well and produced predictions that were very close to the observed
results in the test set \citep{Kapadia2000}, indicating that using such models
justifies further investigation.

@@

\subsection{Artificial Neural Networks}
Artificial neural networks have been widely used in clinical prediction
following their use in diagnosing acute myocardial infarction and various
applications in radiology \citep{Baxt1995}.
One of the earliest applications of neural networks to
length of stay prediction was by Tu \& Guerriere, who used a neural network
with the following structure \citep{Tu1993} to predict length of stay in
intensive care following cardiac surgery:
\begin{figure}[h]
\centering
\includegraphics[scale=0.95]{images/a2-nn}
\caption{A graphical representation of a neural network with one hidden layer,
used by Tu \& Guerriere in predicting LOS.}
\end{figure}

@@

Yang et al. investigated the performance of linear regression, M5 (another
tree-based classification method), and support vector machines in predicting
length of stay for burn patients. The difference between their tree method,
M5, with that of Harper (CART), is that the nodes contain linear regression
functions instead of the usual discrete (two-class) outputs that are used to
successively split the data set. Additionally, they also use support vector
machines (SVMs), perhaps the most powerful classification algorithm in terms of
predictive accuracy \citep{Bellazzi2008}. SVMs work on the principle that when
the data is represented in a high enough dimensional space, a linear hyperplane
can be found where the margin between two different records is as large as
possible. In Figure \ref{fig:svm-example} from Yoo et al. \citep{Yoo2012},
hyperplane 2 is the better plane because its margin is as large as possible. 
\begin{figure}[h]
\label{fig:svm-example}
\centering
\includegraphics[scale=1.05]{images/a2-svm}
\caption{An example of finding the largest margin hyperplane for a two-class
support vector machine classifier.}
\end{figure}
